---
title: "A real-time corpus-based analysis of the discourse around COVID19 in the Australian Twittersphere - Part 4: Period keywords"
author: "Anonymous"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: word_document
bibliography: bibliography.bib
link-citations: yes
---

# Introduction

This document details the data analysis for the project *A real-time corpus-based analysis of the discourse around COVID19 in the Australian Twittersphere*. 

This part of the analysis detects keywords that are significantly associated with specific periods in the covid tweets.

The analysis is initiated by clearing the workspace, setting options, and activating packages as well as functions.

```{r covtwit_04_01, echo=T, eval = T, message=FALSE, warning=FALSE}
# clean current workspace
rm(list=ls(all=T))  
# load libraries
library(tokenizers)
library(tm)
library(stringi)
library(tidytext)
library(tidyverse)
library(SnowballC)
library(here)
# load functions
source(here::here("scripts", "rmwords.R"))
source(here::here("scripts", "CoocStatzFisher.R"))
# load stop words
english_stopwords <- readLines("https://slcladal.github.io/resources/stopwords_en.txt", encoding = "UTF-8")
# set options
options(stringsAsFactors = F)
options(scipen = 999)
options(max.print=10000)
```

## Load data

```{r covtwit_04_02, echo=T, eval = T, message=FALSE, warning=FALSE}
covidtwitterdata <- read.delim(here::here("tables", "covidtwitter2020data_fullyannotated.txt"),
                               comment.char = "", quote = "") %>%
  dplyr::mutate(Period = factor(str_remove_all(Period, "_kw_0")))
# inspect data
head(covidtwitterdata)
```

Summary

```{r covtwit_04_03, echo=T, eval = T, message=FALSE, warning=FALSE}
covidtwitterdata %>%
  dplyr::select(Year, NoTweets, NoElements) %>%
  dplyr::group_by(Year) %>%
  dplyr::summarise(Tweets = n(),
                   Elements = sum(NoElements))
```

Create table with words and frequencies per period

```{r covtwit_04_04, echo=T, eval = T, message=FALSE, warning=FALSE}
perioddata <- covidtwitterdata %>%
  dplyr::group_by(Period) %>%
  unnest_tokens(Word, Text) %>%
  dplyr::group_by(Period, Word) %>%
  dplyr::summarise(Frequency = n()) %>%
  dplyr::group_by(Word) %>%
  dplyr::mutate(Total = sum(Frequency)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(Word = ifelse(Total < 10, "other", Word)) %>%
  dplyr::group_by(Period, Word) %>%
  dplyr::summarise(Frequency = sum(Frequency)) %>%
  tidyr::spread(Period, Frequency) %>%
  replace(is.na(.), 0)
# inspect data
head(perioddata)
```

Function for data processing and extraction of significant keyterms

```{r covtwit_04_05, echo=T, eval = T, message=FALSE, warning=FALSE}
pdperiod <- function(perioddata, Targetperiod){
  periodtb <- perioddata %>%
    tidyr::gather(Period, Frequency, Period1:Period7) %>%
    dplyr::mutate(Period = ifelse(Period != Targetperiod, "Other", Period)) %>%
    dplyr::group_by(Period, Word) %>%
    dplyr::mutate(Frequency = sum(Frequency)) %>%
    unique() %>%
    tidyr::spread(Period, Frequency) %>%
    dplyr::rename(Target = Targetperiod, NonTarget = Other) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(TotalTarget = sum(Target), TotalNonTarget = sum(NonTarget),
                  NRows = n())
  periodkeys <- CoocStatzFisher(periodtb)
  periodkeys <- periodkeys %>%
    dplyr::filter(CorrSignificance != "n.s.") %>%
    dplyr::mutate(Period = rep(Targetperiod, length(CorrSignificance)))
  return(periodkeys)
}
```

# Extract keywords

```{r covtwit_04_06, echo=T, eval = T, message=FALSE, warning=FALSE}
keys_period1 <- pdperiod(perioddata, "Period1")
keys_period2 <- pdperiod(perioddata, "Period2")
keys_period3 <- pdperiod(perioddata, "Period3")
keys_period4 <- pdperiod(perioddata, "Period4")
keys_period5 <- pdperiod(perioddata, "Period5")
keys_period6 <- pdperiod(perioddata, "Period6")
keys_period7 <- pdperiod(perioddata, "Period7")
# inspect data
head(keys_period1)
```

Combine keywords

```{r covtwit_04_07, echo=T, eval = T, message=FALSE, warning=FALSE}
period_keys <- rbind(keys_period1, keys_period2, keys_period3, 
                     keys_period4, keys_period5, keys_period6, keys_period7)
# inspect data
head(period_keys)
```

# Visualize keywords

```{r covtwit_04_08, echo=T, eval = T, message=FALSE, warning=FALSE}
p1d <- period_keys %>%
  dplyr::mutate(Period = as.numeric(stringr::str_remove_all(Period, "Period"))) %>%
  dplyr::filter(CorrSignificance == "p<.001") %>%
  dplyr::filter(Word != "other") %>%
#  dplyr::filter(phi > .001) %>%
  dplyr::mutate(x2 = log(x2)) %>%
  dplyr::mutate(x2 = ifelse(Type == "Overuse", x2, -x2)) 

```


```{r covtwit_04_09, echo=T, eval = T, message=FALSE, warning=FALSE}

p1 <- ggplot(p1d, aes(x = Period, y = x2)) +
#  geom_point()+
  #geom_text(aes(label=Word),hjust=0, vjust=-1,
  #          position=position_jitter(width=.5,height=1), size = 2) +
  geom_text(aes(label = Word), check_overlap = TRUE, vjust = 1.5) +
  labs(x = "Phase", y = "Association strength (logged X2)") +
  scale_x_continuous(breaks = seq(1, 7, 1), 
                     labels= seq(1, 7, 1), 
                     limits = c(0.5, 7.5)) +
  theme_set(theme_bw(base_size = 10)) +
  theme(legend.position="none",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.text.x = element_text(angle = 0, size =7.5),
        plot.margin = unit(c(.2, .2, .2, .2), "cm")) +
  coord_cartesian(ylim = c(-10, 10)) +
  ggsave(file = here::here("images", "Keyterms_Period.png"), 
         height = 4,  width = 6, dpi = 320)
p1
```

Reformat data for mosaicplot

```{r covtwit_04_10, echo=T, eval = T, message=FALSE, warning=FALSE}
topicsperiod <- covidtwitterdata %>%
  dplyr::select(Topic, Period) %>%
  dplyr::mutate(Topic = str_remove_all(Topic, "_.*"),
                Topic = factor(Topic)) %>%
  group_by(Period, Topic) %>%
  dplyr::summarise(Frequency = n()) %>%
  tidyr::spread(Topic, Frequency)
# inspect data
topicsperiod
```


```{r covtwit_04_11, echo=T, eval = T, message=FALSE, warning=FALSE}
topicsperiod_x2 <- chisq.test(as.matrix(topicsperiod[, 2:ncol(topicsperiod)]))
stdres <- topicsperiod_x2$stdres
rownames(stdres) <- topicsperiod$Period
stdres
```


```{r covtwit_04_12, echo=T, eval = T, message=FALSE, warning=FALSE}
topicsperiodgg <- topicsperiod %>%
  gather(Topic, Frequency, Topic1:Topic5) %>%
  cbind(as.vector(stdres)) %>%
  rename(StandardizedResidual = "...4") %>%
  dplyr::group_by(Topic) %>%
  dplyr::mutate(Topic_freq = sum(Frequency),
                Topic_prcnt = Frequency/Topic_freq) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(Period) %>%
  dplyr::mutate(Period_freq = sum(Frequency)) %>%
  dplyr::mutate(Period_prcnt = Frequency/sum(Frequency)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(Tweets_prcnt = Period_freq/sum(unique(Period_freq)))
# inspect results
topicsperiodgg

```



```{r covtwit_04_13, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
topicsperiodgg <- topicsperiodgg %>%
  mutate(StandardizedResidual = case_when(StandardizedResidual <= -4 ~ "min",
                                          StandardizedResidual <= -2 ~ "low",
                                          StandardizedResidual <= 0 ~ "mid",
                                          StandardizedResidual <= 2 ~ "med",
                                          StandardizedResidual <= 4 ~ "high",
                                          TRUE ~ "max")) %>%
  mutate(StandardizedResidual = factor(StandardizedResidual,
                                       levels = c("min", "low", "mid", "med", "high", "max"),
                                       labels = c("<= -4", "<= -2", "<= 0",   "<= 2",  "<= 4", "> 4"))) %>%
  dplyr::mutate(Topic = factor(Topic, labels=c("Topic 1: MEDICAL", "Topic 2: INTERNATIONAL", 
                                               "Topic 3: RESTRICTIONS/HOME", "Topic 4: SPREAD", 
                                               "Topic 5: ECONOMY")))
```



```{r covtwit_04_14, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
topicsperiodgg %>%
  ggplot(aes(Period,
             Topic_prcnt,
             color = Topic, 
             fill = StandardizedResidual,
             width = Period_freq)) + 
  geom_bar(stat = "identity", position = "fill") +
  facet_grid(~Period, scales = "free_x", space = "free_x") +
  scale_fill_brewer(name = "Std. Res.", palette = "RdGy") +
  scale_color_manual(values = rep("black", length(names(table(topicsperiodgg$Topic))))) +
  theme_void()  +
  theme(panel.spacing.x = unit(0, "npc"),
        legend.position = "top",
        legend.key.size = unit(.2, "cm")) +
  ggsave(file = here::here("images", "ggmosaic.png"), 
         height = 4,  width = 12, dpi = 320)
```


## Mosaic plot

```{r covtwit_04_15, echo=T, eval = T, message=FALSE, warning=FALSE}
colnames(topicsperiod) <- c("Period", "Topic1_MEDICAL", "Topic2_INTERNATIONAL", 
                            "Topic3_RESTRICTIONS/HOME", "Topic4_SPREAD", 
                            "Topic5_ECONOMY")
rownames

jpeg(here::here("images", "Covid_Mosaic_PeriodsTopic.jpg"),
    quality = 100, width = 960, height = 1160)
mosaicplot(topicsperiod, shade = TRUE, type = "pearson", main = "", cex = 1.25)
dev.off()
mosaicplot(topicsperiod, shade = TRUE, type = "pearson", main = "")
```

# CFA

Reformat data for CFA

```{r covtwit_04_16, echo=T, eval = T, message=FALSE, warning=FALSE}
topicsperiod_cfa <- covidtwitterdata %>%
  dplyr::select(Topic, Period) %>%
  dplyr::mutate(Topic = str_remove_all(Topic, "_.*"),
                Topic = factor(Topic))
# inspect data
head(topicsperiod_cfa)
```

## Perform CFA

```{r covtwit_04_17, echo=T, eval = T, message=FALSE, warning=FALSE}
library(cfa)
cfa(topicsperiod_cfa)
```

# Outro

```{r}
sessionInfo()
```

