---
title: "A real-time corpus-based analysis of the discourse around COVID19 in the Australian Twittersphere - Part 3: Sentiment analysis"
author: "Anonymous"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: word_document
bibliography: bibliography.bib
link-citations: yes
---

This document details the data analysis for the project *A real-time corpus-based analysis of the discourse around COVID19 in the Australian Twittersphere*. 

This part of the analysis performs a sentiment analysis on the entire data collected in 2020.

The analysis is initiated by clearing the workspace, setting options, and activating packages as well as functions.

```{r covtwit_03_01, echo=T, eval = T, message=FALSE, warning=FALSE}
# clean current workspace
rm(list=ls(all=T))  
# load libraries
library(tokenizers)
library(tm)
library(stringi)
library(tidytext)
library(tidyverse)
library(SnowballC)
library(here)
# set options
options(stringsAsFactors = F)
options(scipen = 999)
options(max.print=10000)
# define image directory
imageDirectory<-"images"
```

Load twitter data

```{r covtwit_03_02, echo=T, eval = T, message=FALSE, warning=FALSE}
covidtwitterdata <- read.delim(here::here("data", "covidtwitter2020data_topics_periods.txt"), comment.char = "", quote = "")
# inspect data
head(covidtwitterdata)
```

# Sentiment analysis

Clean texts for sentiment analysis


Perform sentiment analysis on cleaned texts

```{r covtwit_03_03, echo=T, eval = T, message=FALSE, warning=FALSE}
# load library
library(syuzhet)
# perform sentiment analysis
sentiments <- get_nrc_sentiment(covidtwitterdata$Text)
# add doc_id to sentiments
covidtwitterdata <- cbind(covidtwitterdata, sentiments)
# save data
write.table(covidtwitterdata, here::here("tables", "covidtwitter2020data_fullyannotated.txt"),
            sep = "\t", row.names = F, quote = F)
# inspect data
head(covidtwitterdata)
```

Prepare data for visualizing sentiments per period

```{r covtwit_03_04, echo=T, eval = T, message=FALSE, warning=FALSE}
p8d <- covidtwitterdata %>%
  dplyr::select(Date, Period, Topic, negative, positive) %>%
  dplyr::mutate(Period = str_remove_all(Period, "Period_kw_"),
                Period = as.numeric(Period)) %>%
  dplyr::group_by(Period, Topic) %>%
  dplyr::summarise(positive = sum(positive),
                   negative = sum(negative),
                   Polarity = mean(positive/negative)-1)
# inspect data
head(p8d)
```

```{r covtwit_03_05, echo=T, eval = T, message=FALSE, warning=FALSE}
ggplot(p8d, aes(x = Period, y = Polarity, color = Topic, linetype = Topic)) +
#  facet_wrap(vars(Topic), ncol = 1, scales="free_y") +
#  facet_wrap(vars(Topic), ncol = 1) +
  geom_hline(yintercept = 0, linetype="dashed", color = "red") +
  geom_smooth(se = F) +
  labs(x = "Phase", y = "Polarity\n(negative < 0 | positive > 0)") +
  guides(color=guide_legend(nrow = 3)) +
  theme_set(theme_bw(base_size = 12)) +
  theme(legend.position="top",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.text.x = element_text(angle = 0, size =10)) +
  scale_x_continuous(breaks = seq(1, 7, 1), 
                     labels= seq(1, 7, 1), 
                     limits = c(1, 7)) + 
    scale_y_continuous(limits = c(-.5, .5)) +
    scale_colour_manual(labels = names(table(p8d$Topic)),
                      breaks = names(table(p8d$Topic)),
                      values = paste0("gray", seq(20, 80, 15)), 
                      name = "") +
  scale_linetype_manual(labels = names(table(p8d$Topic)),
                        breaks = names(table(p8d$Topic)),
                        values = seq(1, 5, 1), 
                        name = "") +
  ggsave(file = here::here("images", "Covid_Phases_Polarity.png"), 
         height = 6,  width = 6, dpi = 320)
```


```{r covtwit_03_06, echo=T, eval = T, message=FALSE, warning=FALSE}
p9d <- covidtwitterdata %>%
  dplyr::select(Period, Topic, anger, fear, sadness, anticipation, joy, trust) %>%
  gather(Emotion, Rate, anger:trust) %>%
  dplyr::mutate(Period = str_remove_all(Period, "Period_kw_"),
                Period = as.numeric(Period)) %>%
  dplyr::group_by(Period, Topic, Emotion) %>%
  dplyr::summarise(Rate = mean(Rate)) %>%
  
  dplyr::ungroup() %>%
  dplyr::mutate(Emotion = toupper(Emotion),
                Emotion = factor(Emotion, levels = c("ANGER", "ANTICIPATION",  "FEAR",
                                                     "JOY", "SADNESS", "TRUST")))
# inspect data
head(p9d)
```


```{r covtwit_03_07, echo=T, eval = T, message=FALSE, warning=FALSE}
ggplot(p9d, aes(x = Period, y = Rate, group = Emotion, color = Emotion, linetype = Emotion)) +
#  facet_wrap(vars(Topic), ncol = 2, scales="free_y") +
  facet_wrap(vars(Topic), ncol = 1) +
  geom_line(size = .75) +  
  labs(x = "Phase", y = "Rate (among COVID-19 tweets)") +
  theme_set(theme_bw(base_size = 12)) +
  theme(legend.position="top",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.text.x = element_text(angle = 0, size =10)) +
  scale_colour_manual(values = paste0("gray", rep(c(30, 70), each = 3)),
                      name = "Core emotion",
                      breaks = names(table(p9d$Emotion)),  
                      labels = names(table(p9d$Emotion))) +
  scale_linetype_manual(labels = names(table(p9d$Emotion)),
                        breaks = names(table(p9d$Emotion)),
                        values = rep(c(1:3), 2), 
                        name = "Core emotion") +
  scale_x_continuous(breaks = seq(1, 7, 1), 
                     labels= seq(1, 7, 1), 
                     limits = c(1, 7)) +  
  theme(legend.position="top", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.text.x = element_text(angle = 0, hjust = 1)) +
  ggsave(file = here::here("images", "Covid_Phases_Emotions.png"), 
         height = 6,  width = 5, dpi = 320)
```


```{r covtwit_03_08, echo=T, eval = T, message=FALSE, warning=FALSE}
ggplot(p9d, aes(x = Period, y = Rate, group = Topic, color = Topic, linetype = Topic)) +
#  facet_wrap(vars(Topic), ncol = 2, scales="free_y") +
  facet_wrap(vars(Emotion), ncol = 2) +
  geom_line(size = .75) +  
  labs(x = "Phase", y = "Rate (among COVID-19 tweets)") +
  guides(color=guide_legend(nrow = 3)) +
  theme_set(theme_bw(base_size = 12)) +
  theme(legend.position="top",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.text.x = element_text(angle = 0, size =10)) +
    scale_colour_manual(labels = names(table(p9d$Topic)),
                      breaks = names(table(p9d$Topic)),
                      values = paste0("gray", seq(20, 80, 15)), 
                      name = "") +
  scale_linetype_manual(labels = names(table(p9d$Topic)),
                        breaks = names(table(p9d$Topic)),
                        values = seq(1, 5, 1), 
                        name = "") +
  scale_x_continuous(breaks = seq(1, 7, 1), 
                     labels= seq(1, 7, 1), 
                     limits = c(1, 7)) +  
  theme(legend.position="top", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.text.x = element_text(angle = 0, hjust = 1)) +
  ggsave(file = here::here("images", "Covid_Phases_ByEmotion.png"), 
         height = 6,  width = 6, dpi = 320)
```

Outro

```{r}
sessionInfo()
```

